Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2021-01-06 jss/EpiCompare-paper/EpiCompare-paper.Rmd
Now with a changed intro paragraph.

This paper is broken up into the following sections, section \ref{sec:time-invariant} motivates and showcases tools of time-invariant analysis,  section \ref{sec:overview} prxesents an outline of how \pkg{EpiCompare} aids a practitioner in every step of the pipeline and section \ref{sec:tour} provides a thorough demonstrating of the tools through a detailed example of a full data analysis pipeline.

@@

Although epidemics are by definition a process that evolves over time, epidemics, including the recent COVID-19, often need to be compared in a time-invariant way to understand the processes at play. Additionally, many tools to examine the quantity of the population in each state along the infection process (for example: quantity of susceptible vs infected vs recovered individuals) don't always as intelligently capture the natural connections between the proportion of individuals in these states. Tools in \pkg{EpiCompare} attempt to give the user the ability to extend their toolkit to evaluate epidemics to also include time-invariant approaches and in this section we present benefits of the time-invariant analysis, with 1) motivation for time-invariant analysis through $R_0$ (the initial reproduction number), 2) time-invariant visualization tools for 3-state models (e.g. SIR models), and 3) potential for similar analysis for models with more states (e.g. SEIR models). \ben{Will need to rewrite this section "introduction".}

>>>>

This paper is broken up into the following sections, section \ref{sec:time-invariant} motivates and showcases tools of time-invariant analysis,  section \ref{sec:overview} presents an outline of how \pkg{EpiCompare} aids a practitioner in every step of the pipeline and section \ref{sec:tour} provides a thorough demonstrating of the tools through a detailed example of a full data analysis pipeline.

@@

Epidemics can be difficult to compare to one another due to differences in diseases, locations, or population behaviors, or times.  We focus on comparisons between epidemics adjusting for time, both the relative unit of rate of infection (e.g. days, months, years) and the relative beginning and end of an epidemic.  By adjusting for the unit of infection rate, we can focus on the ``lifetime'' of an epidemic, a view that is concerned more with the total number of lives affected than due to a time constraint.  By adjusting for the relative beginning/end of an infection we can compare, for example the first part of infection from one city to another where this is some time gap in spread.

One way to adjust for both facets of time may seem obvious -- scale the time between zero and one, based on the initial time of infection $t_0$ and the final time $t_F$.  Upon inspection, this method of adjustment requires clear definition of what a beginning and end of an epidemic constitute which requires making one or more choices.  We propose the use of time-invariant analysis as it appears in \pkg{EpiCompare}, which avoids the need of making such a choice.

### 2021-01-06 jss/EpiCompare-paper/EpiCompare-paper.tex
Now with a changed intro paragraph.

Although epidemics are by definition a process that evolves over time,
epidemics, including the recent COVID-19, often need to be compared in a
time-invariant way to understand the processes at play. Additionally,
many tools to examine the quantity of the population in each state along
the infection process (for example: quantity of susceptible vs infected
vs recovered individuals) don't always as intelligently capture the
natural connections between the proportion of individuals in these
states. Tools in \pkg{EpiCompare} attempt to give the user the ability
to extend their toolkit to evaluate epidemics to also include
time-invariant approaches and in this section we present benefits of the
time-invariant analysis, with 1) motivation for time-invariant analysis
through \(R_0\) (the initial reproduction number), 2) time-invariant
visualization tools for 3-state models (e.g.~SIR models), and 3)
potential for similar analysis for models with more states (e.g.~SEIR
models).
\textcolor{violet}{Will need to rewrite this section "introduction".}

>>>>

Epidemics can be difficult to compare to one another due to differences
in diseases, locations, or population behaviors, or times. We focus on
comparisons between epidemics adjusting for time, both the relative unit
of rate of infection (e.g.~days, months, years) and the relative
beginning and end of an epidemic. By adjusting for the unit of infection
rate, we can focus on the ``lifetime'\,' of an epidemic, a view that is
concerned more with the total number of lives affected than due to a
time constraint. By adjusting for the relative beginning/end of an
infection we can compare, for example the first part of infection from
one city to another where this is some time gap in spread.

One way to adjust for both facets of time may seem obvious -- scale the
time between zero and one, based on the initial time of infection
\(t_0\) and the final time \(t_F\). Upon inspection, this method of
adjustment requires clear definition of what a beginning and end of an
epidemic constitute which requires making one or more choices. We
propose the use of time-invariant analysis as it appears in
\pkg{EpiCompare}, which avoids the need of making such a choice.
